---
title: "Wilcoxon Rank Sum Demonstration"
output: 
  html_document:
    code_folding: hide
    theme: cerulean
---

```{r load_library, include=FALSE}
# Use this R-Chunk to load your libraries!
library(tidyverse)
library(DT)
library(pander)
library(readr)
library(car)

```


## Small Sample Size - Same Distribution

```{r message=FALSE, warning=FALSE}
n <- 6                      # sample size

mean <-  20                 # distribution mean

sigma <- 6                  # distribution standard deviation

A <- rnorm(n, mean, sigma) # normally distributed data
B <- rnorm(n, mean, sigma) # normally distributed data

dat <- data.frame(group = c(rep("A", n), rep("B", n)), value = c(A, B))

alpha = 0.05

myTest <- wilcox.test(A, B, paired = FALSE, mu = 0, alternative = "two.sided", conf.level = 1 - alpha)

layout(matrix(c(1,2), 2, 1, byrow = TRUE), heights=c(4,2))
par(mai=c(.5, .5, .6, .01))
colors = c("steelblue", "sienna1")

stripchart(value ~ group, data = dat, pch = 16, col = colors, ylim = c(0,10), xlim = c(0, 40), main = paste("Small Sample - Same Distribution (p-value: ", round(myTest$p.value, 4), ")", ifelse(myTest$p.value < alpha, "\n -- TYPE I ERROR --", "")))

curve(100 * dnorm(x, mean, sigma) + 3, add = TRUE, from=-20, to=40)
abline(h = 3)
lines(c(mean, mean), c(dnorm(mean, mean, sigma)*100 + 3, 3))
boxplot(A, B,
        names = c("A", "B"),
        ylim = c(0, 40),
       # col = ifelse(myTest$p.value < alpha, "red", "skyblue"),
        col = colors,
        boxwex=.5,
        horizontal = TRUE)


pander(myTest)
```


## Small Sample Size - Different Distribution

```{r, echo=TRUE, message=FALSE, warning=FALSE}
n <- 6                      # sample size

mean_A <-  15                 # distribution A mean

mean_B <-  25                 # distribution B mean

sigma <- 6                  # distribution standard deviation

A <- rnorm(n, mean_A, sigma) # normally distributed data
B <- rnorm(n, mean_B, sigma) # normally distributed data

dat <- data.frame(group = c(rep("A", n), rep("B", n)), value = c(A, B))

alpha = 0.05

myTest <- wilcox.test(A, B, paired = FALSE, mu = 0, alternative = "two.sided", conf.level = 1 - alpha)

layout(matrix(c(1,2), 2, 1, byrow = TRUE), heights=c(4,2))
par(mai=c(.5, .5, .6, .01))
colors = c("steelblue", "sienna1")

stripchart(value ~ group, 
           data = dat, 
           pch = 16, 
           col = colors, 
           ylim = c(0,10), 
           xlim = c(0, 40), 
           main = paste("Small Sample - Different Distribution (p-value: ", round(myTest$p.value, 4), ")", ifelse(myTest$p.value > alpha, "\n -- TYPE II ERROR --", "")))

curve(100 * dnorm(x, mean_A, sigma) + 3, add = TRUE, from=-20, to=40, col = colors[1])
curve(100 * dnorm(x, mean_B, sigma) + 3, add = TRUE, from=-20, to=40, col = colors[2])
abline(h = 3)
lines(c(mean_A, mean_A), c(dnorm(mean_A, mean_A, sigma)*100 + 3, 3), col = colors[1])
lines(c(mean_B, mean_B), c(dnorm(mean_B, mean_B, sigma)*100 + 3, 3), col = colors[2])

boxplot(A, B,
        names = c("A", "B"),
        ylim = c(0, 40),
       # col = ifelse(myTest$p.value < alpha, "red", "skyblue"),
        col = colors,
        boxwex=.5,
        horizontal = TRUE)


pander(myTest)

```


## Large Sample Size - Same Distribution

```{r, echo=TRUE, message=FALSE, warning=FALSE}
n <- 30                      # sample size

mean <-  20                 # distribution mean

sigma <- 6                  # distribution standard deviation

A <- rnorm(n, mean, sigma) # normally distributed data
B <- rnorm(n, mean, sigma) # normally distributed data

dat <- data.frame(group = c(rep("A", n), rep("B", n)), value = c(A, B))

alpha = 0.05

myTest <- wilcox.test(A, B, paired = FALSE, mu = 0, alternative = "two.sided", conf.level = 1 - alpha)

layout(matrix(c(1,2), 2, 1, byrow = TRUE), heights=c(4,2))
par(mai=c(.5, .5, .6, .01))
colors = c("steelblue", "sienna1")

stripchart(value ~ group, data = dat, pch = 16, col = colors, ylim = c(0,10), xlim = c(0, 40), main = paste("Large Sample - Same Distribution (p-value: ", round(myTest$p.value, 4), ")", ifelse(myTest$p.value < alpha, "\n -- TYPE I ERROR --", "")))

curve(100 * dnorm(x, mean, sigma) + 3, add = TRUE, from=-20, to=40)
abline(h = 3)
lines(c(mean, mean), c(dnorm(mean, mean, sigma)*100 + 3, 3))
boxplot(A, B,
        names = c("A", "B"),
        ylim = c(0, 40),
       # col = ifelse(myTest$p.value < alpha, "red", "skyblue"),
        col = colors,
        boxwex=.5,
        horizontal = TRUE)

pander(myTest)

```


## Large Sample Size - Different Distribution

```{r, echo=TRUE, message=FALSE, warning=FALSE}
n <- 30                      # sample size

mean_A <-  15                 # distribution A mean

mean_B <-  25                 # distribution B mean

sigma <- 6                  # distribution standard deviation

A <- rnorm(n, mean_A, sigma) # normally distributed data
B <- rnorm(n, mean_B, sigma) # normally distributed data

dat <- data.frame(group = c(rep("A", n), rep("B", n)), value = c(A, B))

alpha = 0.05

myTest <- wilcox.test(A, B, paired = FALSE, mu = 0, alternative = "two.sided", conf.level = 1 - alpha)

layout(matrix(c(1,2), 2, 1, byrow = TRUE), heights=c(4,2))
par(mai=c(.5, .5, .6, .01))
colors = c("steelblue", "sienna1")

stripchart(value ~ group, 
           data = dat, 
           pch = 16, 
           col = colors, 
           ylim = c(0,10), 
           xlim = c(0, 40), 
           main = paste("Large Sample - Different Distribution (p-value: ", round(myTest$p.value, 4), ")", ifelse(myTest$p.value > alpha, "\n -- TYPE II ERROR --", "")))

curve(100 * dnorm(x, mean_A, sigma) + 3, add = TRUE, from=-20, to=40, col = colors[1])
curve(100 * dnorm(x, mean_B, sigma) + 3, add = TRUE, from=-20, to=40, col = colors[2])
abline(h = 3)
lines(c(mean_A, mean_A), c(dnorm(mean_A, mean_A, sigma)*100 + 3, 3), col = colors[1])
lines(c(mean_B, mean_B), c(dnorm(mean_B, mean_B, sigma)*100 + 3, 3), col = colors[2])

boxplot(A, B,
        names = c("A", "B"),
        ylim = c(0, 40),
       # col = ifelse(myTest$p.value < alpha, "red", "skyblue"),
        col = colors,
        boxwex=.5,
        horizontal = TRUE)

pander(myTest)

```


## Non Normal Distributions - Samll Sample - Same Distribution

```{r, echo=TRUE, message=FALSE, warning=FALSE}
n <- 6                      # sample size

df_A <-  3                 # distribution A mean

df_B <-  3                 # distribution B mean

A <- rchisq(n, df_A) # chisq distributed data
B <- rchisq(n, df_B) # chisq distributed data

dat <- data.frame(group = c(rep("A", n), rep("B", n)), value = c(A, B))

alpha = 0.05

myTest <- wilcox.test(A, B, paired = FALSE, mu = 0, alternative = "two.sided", conf.level = 1 - alpha)

layout(matrix(c(1,2), 2, 1, byrow = TRUE), heights=c(4,2))
par(mai=c(.5, .5, .6, .01))
colors = c("steelblue", "sienna1")

stripchart(value ~ group, 
           data = dat, 
           pch = 16, 
           col = colors, 
           ylim = c(0,10), 
           xlim = c(0, 15), 
           main = paste("Small Sample - Same Distribution (p-value: ", round(myTest$p.value, 4), ")", ifelse(myTest$p.value < alpha, "\n -- TYPE I ERROR --", "")))

curve(15 * dchisq(x, df_A) + 3, add = TRUE, from=0, to=20)
curve(15 * dchisq(x, df_B) + 3, add = TRUE, from=0, to=20)
abline(h = 3)
lines(c(df_A, df_A), c(dchisq(df_A, df_A)*15 + 3, 3))
lines(c(df_B, df_B), c(dchisq(df_B, df_B)*15 + 3, 3))

boxplot(A, B,
        names = c("A", "B"),
        ylim = c(0, 15),
       # col = ifelse(myTest$p.value < alpha, "red", "skyblue"),
        col = colors,
        boxwex=.5,
        horizontal = TRUE)


pander(myTest)

```


## Non Normal Distributions - Small Sample - Different Distribution

```{r, echo=TRUE, message=FALSE, warning=FALSE}
n <- 6                      # sample size

df_A <-  3                 # distribution A mean

df_B <-  6                 # distribution B mean

A <- rchisq(n, df_A) # chisq distributed data
B <- rchisq(n, df_B) # chisq distributed data

dat <- data.frame(group = c(rep("A", n), rep("B", n)), value = c(A, B))

alpha = 0.05

myTest <- wilcox.test(A, B, paired = FALSE, mu = 0, alternative = "two.sided", conf.level = 1 - alpha)

layout(matrix(c(1,2), 2, 1, byrow = TRUE), heights=c(4,2))
par(mai=c(.5, .5, .6, .01))
colors = c("steelblue", "sienna1")

stripchart(value ~ group, 
           data = dat, 
           pch = 16, 
           col = colors, 
           ylim = c(0,10), 
           xlim = c(0, 15), 
           main = paste("Small Sample - Different Distribution (p-value: ", round(myTest$p.value, 4), ")", ifelse(myTest$p.value > alpha, "\n -- TYPE II ERROR --", "")))

curve(15 * dchisq(x, df_A) + 3, add = TRUE, from=0, to=20, col = colors[1])
curve(15 * dchisq(x, df_B) + 3, add = TRUE, from=0, to=20, col = colors[2])
abline(h = 3)
lines(c(df_A, df_A), c(dchisq(df_A, df_A)*15 + 3, 3), col = colors[1])
lines(c(df_B, df_B), c(dchisq(df_B, df_B)*15 + 3, 3), col = colors[2])

boxplot(A, B,
        names = c("A", "B"),
        ylim = c(0, 15),
       # col = ifelse(myTest$p.value < alpha, "red", "skyblue"),
        col = colors,
        boxwex=.5,
        horizontal = TRUE)


pander(myTest)

```



## Non Normal Distributions - Large Sample - Same Distribution

```{r, echo=TRUE, message=FALSE, warning=FALSE}
n <- 30                      # sample size

df_A <-  3                 # distribution A mean

df_B <-  3                 # distribution B mean

A <- rchisq(n, df_A) # chisq distributed data
B <- rchisq(n, df_B) # chisq distributed data

dat <- data.frame(group = c(rep("A", n), rep("B", n)), value = c(A, B))

alpha = 0.05

myTest <- wilcox.test(A, B, paired = FALSE, mu = 0, alternative = "two.sided", conf.level = 1 - alpha)

layout(matrix(c(1,2), 2, 1, byrow = TRUE), heights=c(4,2))
par(mai=c(.5, .5, .6, .01))
colors = c("steelblue", "sienna1")

stripchart(value ~ group, 
           data = dat, 
           pch = 16, 
           col = colors, 
           ylim = c(0,10), 
           xlim = c(0, 15), 
           main = paste("Large Sample - Same Distribution (p-value: ", round(myTest$p.value, 4), ")", ifelse(myTest$p.value < alpha, "\n -- TYPE I ERROR --", "")))

curve(15 * dchisq(x, df_A) + 3, add = TRUE, from=0, to=20)
curve(15 * dchisq(x, df_B) + 3, add = TRUE, from=0, to=20)
abline(h = 3)
lines(c(df_A, df_A), c(dchisq(df_A, df_A)*15 + 3, 3))
lines(c(df_B, df_B), c(dchisq(df_B, df_B)*15 + 3, 3))

boxplot(A, B,
        names = c("A", "B"),
        ylim = c(0, 15),
       # col = ifelse(myTest$p.value < alpha, "red", "skyblue"),
        col = colors,
        boxwex=.5,
        horizontal = TRUE)


pander(myTest)

```


## Non Normal Distributions - Large Sample - Different Distribution

```{r, echo=TRUE, message=FALSE, warning=FALSE}
n <- 30                      # sample size

df_A <-  3                 # distribution A mean

df_B <-  6                 # distribution B mean

A <- rchisq(n, df_A) # chisq distributed data
B <- rchisq(n, df_B) # chisq distributed data

dat <- data.frame(group = c(rep("A", n), rep("B", n)), value = c(A, B))

alpha = 0.05

myTest <- wilcox.test(A, B, paired = FALSE, mu = 0, alternative = "two.sided", conf.level = 1 - alpha)

layout(matrix(c(1,2), 2, 1, byrow = TRUE), heights=c(4,2))
par(mai=c(.5, .5, .6, .01))
colors = c("steelblue", "sienna1")

stripchart(value ~ group, 
           data = dat, 
           pch = 16, 
           col = colors, 
           ylim = c(0,10), 
           xlim = c(0, 15), 
           main = paste("Large Sample - Different Distribution (p-value: ", round(myTest$p.value, 4), ")", ifelse(myTest$p.value > alpha, "\n -- TYPE II ERROR --", "")))

curve(15 * dchisq(x, df_A) + 3, add = TRUE, from=0, to=20, col = colors[1])
curve(15 * dchisq(x, df_B) + 3, add = TRUE, from=0, to=20, col = colors[2])
abline(h = 3)
lines(c(df_A, df_A), c(dchisq(df_A, df_A)*15 + 3, 3), col = colors[1])
lines(c(df_B, df_B), c(dchisq(df_B, df_B)*15 + 3, 3), col = colors[2])

boxplot(A, B,
        names = c("A", "B"),
        ylim = c(0, 15),
       # col = ifelse(myTest$p.value < alpha, "red", "skyblue"),
        col = colors,
        boxwex=.5,
        horizontal = TRUE)


pander(myTest)

```
